---
title: "Getting Started with nflendzoneModel"
output: html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with nflendzoneModel}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

`nflendzoneModel` provides pre-compiled CmdStan models for Bayesian analysis of NFL game data. This vignette demonstrates the basic workflow.

## Prerequisites

Before using this package, ensure CmdStan is installed:

```{r eval = FALSE}
# Check if CmdStan exists
instantiate::stan_cmdstan_exists()

# If not, install it
cmdstanr::install_cmdstan()
```

## Loading the Package

```{r setup, eval = FALSE}
library(nflendzoneModel)
library(cmdstanr)
library(posterior)
```

## Accessing Models

The package provides two pre-compiled Stan models:

```{r eval = FALSE}
# Get the fit model
fit_mod <- get_team_strength_fit_model()

# Get the generated quantities model
gq_mod <- get_team_strength_gq_model()
```

These models are already compiled and ready to use.

## Model Structure

### Fit Model

The team strength fit model is a hierarchical state-space model with:

- **Team strengths**: AR(1) process across weeks
- **Home field advantage**: League-level + team-specific components
- **Seasonal transitions**: Carryover effects between seasons

### Generated Quantities Model

The GQ model takes a fitted model and generates predictions for out-of-sample games.

## Basic Workflow

```{r eval = FALSE}
# 1. Prepare your data (you need to create these helpers)
# fit_data <- create_stan_data(...)

# 2. Fit the model
# fit <- fit_team_strength_model(
#   stan_data = fit_data,
#   chains = 4,
#   parallel_chains = 4,
#   iter_warmup = 1000,
#   iter_sampling = 2000
# )

# 3. Generate predictions
# gq_data <- prepare_gq_data(...)
# predictions <- generate_team_predictions(
#   fit = fit,
#   gq_data = gq_data
# )

# 4. Extract results
# draws <- predictions$draws()
# summary <- predictions$summary()
```

## Next Steps

- See the model specifications in `src/stan/`
- Develop data preparation helpers for your NFL data
- Extend with custom analysis functions