---
title: "Getting Started with nflendzoneModel"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with nflendzoneModel}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

`nflendzoneModel` provides pre-compiled CmdStan models for Bayesian analysis of NFL game data. This vignette demonstrates the basic workflow.

## Prerequisites

Before using this package, ensure CmdStan is installed:

```{r eval = FALSE}
# Check if CmdStan exists
instantiate::stan_cmdstan_exists()

# If not, install it
cmdstanr::install_cmdstan()
```

## Loading the Package

```{r setup, eval = FALSE}
library(nflendzoneModel)
library(cmdstanr)
library(posterior)
```

## Accessing Models

The package exposes pre-compiled CmdStan executables. Retrieve them with `instantiate::stan_package_model()` when you need direct access:

```{r eval = FALSE}
fit_mod <- instantiate::stan_package_model(
  name = "team_strength_fit",
  package = "nflendzoneModel"
)

gq_mod <- instantiate::stan_package_model(
  name = "team_strength_gq",
  package = "nflendzoneModel"
)
```

These objects point to the compiled binaries and can be reused across sessions once CmdStan has been configured.

## Model Structure

### Fit Model

The team strength fit model is a hierarchical state-space model with:

- **Team strengths**: AR(1) process across weeks
- **Home field advantage**: League-level + team-specific components
- **Seasonal transitions**: Carryover effects between seasons

### Generated Quantities Model

The GQ model takes a fitted model and generates predictions for out-of-sample games.

## Basic Workflow

```{r eval = FALSE}
# 1. Prepare Stan-ready data (see the repository's inst/examples scripts for reference)
# fit_data <- build_team_strength_data(...)

# 2. Fit the model using the packaged helper
fit <- fit_team_strength_model(
  stan_data = fit_data,
  chains = 4,
  parallel_chains = 4,
  iter_warmup = 1000,
  iter_sampling = 2000
)

# 3. Assemble generated-quantities input for the games you want to score
# gq_data <- build_gq_data(fit_data, future_schedule, ...)

# 4. Generate predictions from the fitted draws
predictions <- generate_team_predictions(
  fit = fit,
  gq_data = gq_data,
  parallel_chains = 4
)

# 5. Work with posterior draws or summaries
draws <- predictions$draws()
posterior::summarise_draws(draws, c("mean", "sd"))
```

## Next Steps

- See the model specifications in `src/stan/`
- Develop data preparation helpers for your NFL data
- Extend with custom analysis functions